---
title: "Oxygen & Voltage Gradients"
output: 
  html_document:
    code_folding: hide
    toc: yes
    toc_float: yes
    df_print: paged
---

# Introduction

This rmd. file imports oxygen and voltage data from Google sheet
Data is used to plot oxygen and voltage gradient maps of the winogradsky column, exponential and linear models were used to analyze data

## Loading Packages 

```{r loading packages}
library(tidyverse) 
library(googledrive) 
library(googlesheets4)
#library(ggplot2)
library(knitr)
library(minpack.lm)
library(broom)
library(purrr)
library(lme4)
library(patchwork)
```

## Importing from google sheet

```{r load CleanOxygenData direct from googlesheet}

googlesheets4::gs4_deauth()

CleanData_Oxygen <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1hNakCyqPsTNRFpf1lGjoBQ-papyZ6xHKT6mcvh3LMb8/edit?pli=1&gid=842906991#gid=842906991", "CleanData_Oxygen") |>
  mutate(Sample = as.character(Sample))

kable(CleanData_Oxygen) 
```

```{r load CleanVoltageData direct from googlesheet}

googlesheets4::gs4_deauth()

CleanData_Voltage <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1hNakCyqPsTNRFpf1lGjoBQ-papyZ6xHKT6mcvh3LMb8/edit?pli=1&gid=842906991#gid=842906991", "CleanData_Voltage") |>
  mutate(Sample = as.character(Sample))

kable(CleanData_Voltage)
```

Combine O2 & Voltage data by sample & depth
```{r gradients_df}
Gradients <- left_join(x = CleanData_Oxygen, y = CleanData_Voltage, by = c("Sample", "Depth_cm"))
```


## Oxygen Gradient Plot
```{r Oxygen gradients}
ggplot(data = CleanData_Oxygen) +
geom_point(aes(y = DO_µg_L, x = Depth_cm)) + 
  #scale_y_reverse() +
  labs( title= "Oxygen Gradients", caption= "Figure 1. Dissolved Oxygen (DO) (µg/L) measured in four Winogradsky colums at 5 depths (cm)") +
  #theme(plot.caption= element_text(size = 12, hjust=5)) +
  #geom_path(aes(y = Depth_cm, x = DO_µg_L)) +
  facet_grid(cols = vars(Sample)) +
  theme_bw()
#+ scale_y_continuous(limits = c(0, 300))
```
## Fitting an Exponential Decay Model
```{r decay fits}
#define exponential decay function for data fitting.
exp_decay <- function(x, i, mu){y = i * exp(mu * x)}


O2_nest <- CleanData_Oxygen |>  #alternate forward pipe is %>% loaded with tidyverse
  nest(.by = "Sample") |>
  mutate(DecayFit = purrr::map(data, ~nlsLM(DO_µg_L ~ exp_decay(x = Depth_cm, i, mu),
                                            data = .x)),
         DecayTidy = purrr::map(DecayFit, tidy),
         DecayParam = purrr::map(DecayFit, glance),
         DecayPredict = purrr::map(DecayFit, augment)
         )

 
```


```{r plot O2 decay fits}
O2_nest |>
  unnest(cols = c(DecayPredict)) |>
  ggplot() +
  geom_point(aes(x = Depth_cm, y = DO_µg_L)) +
  geom_line(aes(x = Depth_cm, y = .fitted)) +
  geom_point(aes(x = Depth_cm, y = .resid), colour = "red") +
  labs( title= "Oxygen Decay Fits", caption= "Figure 2. Dissolved Oxygen (DO) (µg/L) measured in four Winogradsky colums at 5 depths (cm), fitted to an exponential decay model with residuals") +
  facet_grid(cols = vars(unlist(Sample))) +
  theme_bw()
```

```{r show fit parameters}
O2_nest |>
unnest(cols = c(DecayTidy)) |>
 select(-c(data, DecayFit, DecayParam, DecayPredict)) |>
  select(-c(statistic)) |>
  pivot_wider(id_cols = Sample, names_from = term, values_from = c(estimate, std.error, p.value)) |>
  kable(digits=4)
```

## Voltage Gradient Plot
```{r Plotting Voltage Gradients in water at 10 seconds}

ggplot(data = Gradients) +
  geom_point(aes(x = Depth_cm, y = Voltage_Water_10s_mv)) + 
  labs( title= "Voltage Gradients", caption= "Figure 3. Voltage in water at 10 seconds (mv) measured in four Winogradsky colums at 5 depths (cm)") +
  facet_grid(cols = vars(Sample)) + 
  theme_bw()
```

## Fitting a Linear Model
```{r Plotting With a linear fit}

  Linear_fit <- Gradients %>% 
  group_by(Sample) %>%
  group_modify(~ {

  linear <- lm(Voltage_Water_10s_mv ~ Depth_cm, data = .x, na.action = na.exclude)
  augment(linear, data = .x)}) %>%
  ungroup()


  Linear_fit |>   
    ggplot(aes(x = Depth_cm, y = Voltage_Water_10s_mv)) +
    geom_point() +
    geom_line(aes(y = .fitted)) +
    #Showing residuals
    #geom_point(aes(x = Depth_cm, y = .resid), colour = "red") +
    geom_segment(aes(xend = Depth_cm, yend = .fitted), color = "red") +
    labs( title= "Voltage Gradient Fits", caption= "Figure 4. Voltage in water at 10 seconds (mv) measured in four Winogradsky colums at 5 depths (cm), fitted to a linear model with residuals") +
    facet_grid(cols = vars(unlist(Sample))) +
    theme_bw()


```


```{r Examining linear fit parameters}
 
Linear_fit_param <- Linear_fit |>
  group_by(Sample) |>
  group_modify(~ {
    linear_model <- lm(Voltage_Water_10s_mv ~ Depth_cm, data = .x)
    tidy_model <- tidy(linear_model)  
    tidy_model$Sample <- unique(.x$Sample)  # Ensure Sample column is added
    return(tidy_model)  
  }) |>
  ungroup() 

Linear_fit_param <- Linear_fit_param |>
  mutate(term = str_replace(term, "\\(Intercept\\)", "I"))

Linear_fit_param |>
  select(-c(statistic)) |>
  pivot_wider(id_cols = Sample, names_from = term, values_from = c(estimate, std.error, p.value)) |>
  kable(digits=4)
 
```

## Oxygen and Voltage Gradients
```{r oxygen and voltage gradients}

    Oxygen <- ggplot(data = Gradients) +
      geom_point(aes(x = Depth_cm, y = DO_µg_L)) +
   labs( title= "Oxygen Gradient Comparison", caption= "Figure 5. Dissolved Oxygen (DO) (µg/L) measured in four Winogradsky colums at 5 depths (cm)") +
      facet_grid(cols = vars(Sample)) + 
      theme_bw()

    Voltage <- ggplot(data = Gradients) +
      geom_point(aes(x = Depth_cm, y = Voltage_Water_10s_mv)) +
      labs( title= "Voltage Gradient Comparison", caption= "Figure 6. Voltage in water at 10 seconds (mv) measured in four Winogradsky colums at 5 depths (cm)") +
      facet_grid(cols = vars(Sample)) + 
      theme_bw()
      
  print(Oxygen/Voltage)
  
  
```


```{r Oxygen and Voltage relationship}

ggplot(data = Gradients) +
      geom_point(aes(x = DO_µg_L, y = Voltage_Water_10s_mv)) +
      geom_smooth(aes(x = DO_µg_L, y = Voltage_Water_10s_mv), method = "lm") +
  labs( title= "Oxygen and Voltage Gradient Relationship", caption= "Figure 7. Dissolved Oxygen (DO) (µg/L) (x) and Voltage (y) (mv) measured in four Winogradsky colums at 5 depths (cm)") +
      facet_grid(cols = vars(Sample)) + 
      theme_bw()

```

```{r Oxygen and Voltage param}

Oxygen_Voltage <- Gradients %>%
  group_by(Sample) %>%  # Group by Sample
  group_modify(~ {
    model <- lm(Voltage_Water_10s_mv ~ DO_µg_L, data = .x)
    tidy(model) })

Oxygen_Voltage_filtered <- Oxygen_Voltage %>%
  filter(term == "DO_µg_L") %>%  
  select(Sample, estimate, std.error, p.value) 
 
Oxygen_Voltage_filtered %>%
kable(digits = 4)

```



