---
title: "R Notebook"
output: html_notebook
---

# Microcosms in Winogradsky Columns

## Summary

Winogradsky columns are clear cylinders filled with sediment and water (Esteban et al., 2015). When incubated with light, they grow communities of microorganisms with spacial distribution differences (Esteban et al., 2015). Layers in the column are organized by the different metabolic strategies of microorganisms (Esteban et al., 2015). Winogradsky columns are model systems for studying microbial communities. This project seeks to use the columns to investigate the effect of climate change on the microbial community distributions in Sackville, New Brunswick. By collecting sediment from the Sackville Waterfowl park, the effects of salinity and temperature associated with climate change will be monitored over a 4-week period. This will be done visually, and with dissolved oxygen and voltage gradient measurements. It is hypothesized that there will be changes in microbe abundance, diversity, and community composition with climate change, which will be simulated by changes in temperature and salinity. 

## Highly Qualified Personnel

- Daytona Evelyn McMackin, Mount Allison University, demcmackin@mta.ca
- Lindsey Nicole Hood, Mount Allison University, lnhood@mta.ca
- Douglas A. Campbell, Mount Allison University, dcampbel@mta.ca, ORCID 0000-0001-8996-5463

## Principle Investigators

- Daytona Evelyn McMackin, Mount Allison University, demcmackin@mta.ca
- Lindsey Nicole Hood, Mount Allison University, lnhood@mta.ca

## Primary Contacts  

- Daytona Evelyn McMackin, Mount Allison University, demcmackin@mta.ca
- Lindsey Nicole Hood, Mount Allison University, lnhood@mta.ca

## Data sources

- No data was used from external providers

## Funding sources

- Mount Allison University, and Douglas A. Campbell

## Keywords

Winogradsky column, Climate change, Sea level, Temperature, Oxygen gradients, Voltage gradients, Niches 

## Additional information and support

- Sensitive Data Flag - Human Participants:  NO
- Sensitive Data Flag - Indigenous Partnerships: NO
- Sensitive Data Flag - Government Partnerships: NO
- Sensitive Data Flag - Industry Partnerships: NO
- Access Restrictions - None

## Software  

For data management, Google sheets was used.
For photo capture, the camera of a Samsung S23 was used.
For data analysis, RStudio version 2023.06.01 and R version 4.4.1 were used.

## Repo content information

The repository Winogradsky is organized into folders. Within documents, you will find photos and this README file. To find code, there are separate rmd. for each analysis. 

### MetaDataCatalog
URL for MetaDataCatalog, or MetaDataCatalog filename
example:
https://docs.google.com/spreadsheets/d/1rlj03Q_wq83qlyiY7mXq8BK6C-L0Ko799JsyhfA1TGI/edit#gid=0
Use standard variable names for MetaDataCatalog

##Load Packages
```{r packages, include = FALSE}

library(tidyverse) #core packages from Tidyverse
library(googledrive) #access to GoogleDrive
library(googlesheets4)
```

##Importing MetaData
```{r load Catalog direct from googlesheet, results = "hide"}
googlesheets4::gs4_deauth()

MetaDatasheet <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1hNakCyqPsTNRFpf1lGjoBQ-papyZ6xHKT6mcvh3LMb8/edit?pli=1&gid=1390393511#gid=1390393511")
```

### Importing Data Dictionary
```{r load Catalog direct from googlesheet data_dictionary, results = "hide"}
googlesheets4::gs4_deauth()

DataDictionary <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1hNakCyqPsTNRFpf1lGjoBQ-papyZ6xHKT6mcvh3LMb8/edit?pli=1&gid=1422712698#gid=1422712698", "Data_dictionary")
```

### Data/RawData

Raw data files in various formats. Original files generated by analytical equipment, received from a data provider or outside contractor, etc.
Subfolders contain files from a single instrument.
Do not create swarms of superfluous sub-folders.
```{r load Catalog direct from googlesheet endpointrawdata, results = "hide"}

googlesheets4::gs4_deauth()

RawData <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1hNakCyqPsTNRFpf1lGjoBQ-papyZ6xHKT6mcvh3LMb8/edit?pli=1&gid=1422712698#gid=1422712698", "RawData")
```


### Code

In our code folder:
-There is the voltage gradients (Voltage_gradients.rmd & html)
-There is the oxygen gradients (Oxygen_gradients.rmd & html)
-There is the joined gradients of oxygen and voltage (Gradients.rmd & html)
-There is slide photos importing (Slide_Photos.rmd & html)
-There is a importation file (FileImportFunctions.rmd)

### Docs

In our document folder:
-This file (README_Winogradsky.Rmd) that provides an overview of the repository
-Instructions on how the repository is formatted (INSTRUCTIONS.md)
-A document overviewing this folder (docs.md)
-Folders containing slide photos and column photos for qualitative analysis

### Output

Output from knit .Rmd, Figures and tables produced from analysis.
Do not generate swarms of superfluous data exploration figures.

###Discussion

#Voltage and Oxygen
Voltage and oxygen show a relationship in microbial communities (Christwardana et al., 2021). The relationship has been seen to be positively exponentially proportional between the two factors (Christwardana et al., 2021). In samples 1003 and 1004, a positive relationship is seen, unlike in 1001 and 1002.This is due to seawater being a good conductor in comparison to freshwater, increasing the voltage (Christwardana et al., 2021).

#Niches
Over the course of the four week period, the microbial community shifted. This is due to niche differences between different microbes (Esteban et al., 2015). Due to increases in depth, and therefore light, oxygen, and voltage, communities change according to depth (Esteban et al., 2015). Roughly, the surface sees a higher occurrence of cyanobacteria and other phytoplankton that rely on high light levels for photosynthesis. Below that, heterotrophic bacteria use the byproducts of photosynthesis to fuel their metabolism, which causes a chain reaction between layers in the Winogradsky column. As oxygen is finally used up in the bottom layers, sulfate-reducing bacteria take over, surviving anoxic conditions.In 1001 especially, the layers are visible in the high light side. 

#Microbial Diversity under Climate Change
Between our freshwater and seawater conditions, there were some obvious changes.
In the seawater condition, we saw:
-Iron come out of solution 
-A loss a diversity 
-A loss of an active microbial community in the water column
-A darkening of the sediment 

Between the temperature conditions, we saw:
-A decrease in microbial diversity, even in the other freshwater condition
-A lack of cyanobacterial/green photosynthetic growth, as seen with the sole occurrence of diatoms in 1001
-An increase in orange growth, perhaps iron-oxidizing bacteria

Overall, there is a qualitative effect of climate change (sea water rise and temperature increase) on the microbial communities of Sackville's Waterfowl park. These changes are important to take into consideration as the threats of climate change on a flood-vulnerable town like Sackville are very real. 

## Bibliography

Christwardana, M., Yoshi, L. A., Setyonadi, I., & Maulana, M. R. (2021). Correlation between voltage, dissolved oxygen, and power density of yeast microbial fuel cell in different environmental waters as catholyte. 050001. https://doi.org/10.1063/5.0045395

Esteban, D. J., Hysa, B., & Bartow-McKenney, C. (2015). Temporal and Spatial Distribution of the Microbial Community of Winogradsky Columns. PLoS ONE, 10(8), e0134588. https://doi.org/10.1371/journal.pone.0134588
