---
title: "R Notebook"
output: html_notebook
---

# Microcosms in Winogradsky Columns

## Summary

Winogradsky columns are clear cylinders filled with sediment and water (Esteban et al., 2015). When incubated with light, they grow communities of microorganisms with spacial distribution differences (Esteban et al., 2015). Layers in the column are organized by the different metabolic strategies of microorganisms (Esteban et al., 2015). Winogradsky columns are model systems for studying microbial communities. This project seeks to use the columns to investigate the effect of climate change on the microbial community distributions in Sackville, New Brunswick. By collecting sediment from the Sackville Waterfowl park, the effects of salinity and temperature associated with climate change will be monitored over a 4-week period. This will be done visually, and with dissolved oxygen and voltage gradient measurements. It is hypothesized that there will be changes in microbe abundance, diversity, and community composition with climate change. 

## Highly Qualified Personnel

- Daytona Evelyn McMackin, Mount Allison University, demcmackin@mta.ca
- Lindsey Nicole Hood, Mount Allison University, lnhood@mta.ca
- Douglas A. Campbell, Mount Allison University, dcampbel@mta.ca, ORCID 0000-0001-8996-5463

## Principle Investigators

- Daytona Evelyn McMackin, Mount Allison University, demcmackin@mta.ca
- Lindsey Nicole Hood, Mount Allison University, lnhood@mta.ca

## Primary Contacts  

- Daytona Evelyn McMackin, Mount Allison University, demcmackin@mta.ca
- Lindsey Nicole Hood, Mount Allison University, lnhood@mta.ca

## Data sources

- Provide links to any data used from external providers .

## Funding sources

- List your funding sources, grant names, and grant numbers as applicable.

## Keywords

Winogradsky column, Climate change, Sea level, Temperature, Oxygen gradients, Voltage gradients, Niches 

## Additional information and support

- Sensitive Data Flag - Human Participants:  NO
- Sensitive Data Flag - Indigenous Partnerships: NO
- Sensitive Data Flag - Government Partnerships: NO
- Sensitive Data Flag - Industry Partnerships: NO
- Access Restrictions - None

## Software  

For data management, Google sheets was used.
For photo capture, the camera of a Samsung S23 was used.
For data analysis, RStudio version 2023.06.01 and R version 4.4.1 were used.

## Repo content information

Use the space below to summarize the structure of the repository with a description of each folder, as applicable.

### MetaDataCatalog
URL for MetaDataCatalog, or MetaDataCatalog filename
example:
https://docs.google.com/spreadsheets/d/1rlj03Q_wq83qlyiY7mXq8BK6C-L0Ko799JsyhfA1TGI/edit#gid=0
Use standard variable names for MetaDataCatalog
##Load Packages
```{r packages, include = FALSE}

library(tidyverse) #core packages from Tidyverse
library(googledrive) #access to GoogleDrive
library(googlesheets4)
```

##Importing MetaData
```{r load Catalog direct from googlesheet, results = "hide"}
googlesheets4::gs4_deauth()

MetaDatasheet <- googlesheets4::read_sheet("https://docs.google.com/spreadsheets/d/1hNakCyqPsTNRFpf1lGjoBQ-papyZ6xHKT6mcvh3LMb8/edit?pli=1&gid=1390393511#gid=1390393511")
```

### Importing Data Dictionary
```{r load Catalog direct from googlesheet, results = "hide}
googlesheets4::gs4_deauth()

DataDictionary <- googlesheets4::range_read("https://docs.google.com/spreadsheets/d/1hNakCyqPsTNRFpf1lGjoBQ-papyZ6xHKT6mcvh3LMb8/edit?pli=1&gid=1422712698#gid=1422712698")
```

URL for Data Dictionary, or Data Dictionary filename
example:
https://docs.google.com/spreadsheets/d/1KzHZETwASnt4XW69HYufHwlWaE_E-yHkLDqetiOYqBg/edit#gid=671854404

### Data/RawData

Raw data files in various formats. Original files generated by analytical equipment, received from a data provider or outside contractor, etc.
Subfolders contain files from a single instrument.
Do not create swarms of superfluous sub-folders.

### Data/CleanData

Clean data in formats for long-term storage. Modified data with the appropriate column/row headers and data structure.

### Data/ProcessedData

Processed data in formats for long-term storage.

### Code

Scripts for processing raw data into cleaned data, outside derived code, and user derived code.
A folder OldR is used to store outdated code.
Typically organize .Rmd in modules; Import; Tidy; Process saving .Rds out of each step.

### Docs

Project notes, other documentation, etc.

### Output

Output from knit .Rmd, Figures and tables produced from analysis.
Do not generate swarms of superfluous data exploration figures.

## Bibliography
